---
title: "EHDI Website - Figuring out figures - v1"
author: "Veronica Chaitan"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    html_document:
      plots: 
        style: Normal
        align: left
      page_margins:
        bottom: 1
        top: 1
        right: 1
        left: 1
        header: 0.5
        footer: 0.5
        gutter: 0
      page_size:
        width: 8.5
        height: 11
        orient: "portrait"
    fig_caption: TRUE
    df_print: kable
    fig_height: 2.5
    fig_width: 4
always_allow_html: TRUE
editor_options: 
  chunk_output_type: console
---
[//]: # CSS style arguments

<style type="text/css">

@import url("https://fonts.googleapis.com/css?family=Open+Sans:400,400i,700,700i&display=swap");

h1, h2, h3, h4, h5, body, ul, ol, p, td, th {
font-family: 'Open Sans', sans-serif;

}


body{ /* Normal  */
      font-size: 20px;
      counter-reset:table figure;
  }

.table{
  width:auto;
  font-size:12px;
}

td, th{
  padding-left:10px;
  text-align: right;
  font-family: 'Open Sans', sans-serif;
}

caption::before{
  counter-increment: table;
  content: "Table " counter(table) ": ";
}

.caption::before{
  counter-increment: figure;
  content: "Figure " counter(figure) ": ";
}

caption, .caption{
  font-style:italic;
  font-size: 16px;
  margin-top:0.5em;
  margin-bottom:0.5em;
  width:80%;
  text-align: left;
}

#TOC {
  font-size: 17px;
  width: 100%;
}

</style>

```{r setup, include=FALSE}

# You can set a default figure size here so you don't have to customize every time
knitr::opts_chunk$set(fig.width=8, fig.height=6, warning=FALSE, message=FALSE, cache=FALSE, echo=FALSE)

```

```{r}

# '# PROLOG   ##################################################################'
# '#                                                                           #'
# '# PROJECT: GCATS                                                            #'
# '# PURPOSE: To explore 2018 and 2022 GCATS data for menthol report           #'
# '# DIR:     Box\CPHSS\GCATS\Analysis                                         #'
# '#                                                                           #'
# '# DATA:    Box\CPHSS\GCATS\Data\Data_ValueLabels\GCATS18_FullSample.csv     #'
# '#          Box\CPHSS\GCATS\Data\Data_ValueLabels\GCATS22_FullSample.csv     #'
# '# AUTHOR:  Jessica Price & Veronica L Chaitan                               #'
# '# CREATED: Aug 09, 2023                                                     #'
# '# LATEST:  AUG 09, 2023                                                     #'
# '# NOTES:                                                                    #'
# '#                                                                           #'
# '# PROLOG   ##################################################################'

library(crosstalk)
library(plotly)
library(kableExtra)
library(magrittr)
library(tidyverse)

# for NA in kable
options(knitr.kable.NA = "")

# Theme settings
# theme_set(theme_minimal()) - doesn't carry over across chunks in RMD
# In general
mytheme <- theme_minimal() + 
  theme(text = element_text(size=11, color="gray23"),
        panel.grid = element_blank(),
        axis.text.x = element_text(size=8, color="gray23"),
        axis.text.y = element_text(size=10, color="gray23"),
        axis.title = element_blank(),
        plot.title = element_blank(),
        legend.position = "top",
        legend.title = element_blank())

# Read in the latest dataset
r <- read_csv("../website_ready_data/Final_merged_all_data_v7.csv") %>%
  # remove National sum
  filter(State != "National Sum")
  

# Read in the states to get states abbr
states <- read_csv("../website_ready_data/states.csv")

# Order the states
r <- r %>%
  mutate(State = factor(State,
                        levels=c("National Average",sort(unique(states$State))))) 

# Get rid of infinite...
r <- r %>%
  mutate(PerLTF=case_when(is.infinite(PerLTF)~NA_real_,
                          TRUE~PerLTF),
         PerD=case_when(is.infinite(PerD)~NA_real_,
                          TRUE~PerD))


# State codes
r <- r %>%
  # Joining with `by = join_by(State)`
  left_join(states)



# Account for missing states as NA?

# Make a tibble with every state and year
stateyear <- tibble(State=rep(unique(r$State),7),
                    Year=c(rep(2014,52),rep(2015,52),rep(2016,52),
                           rep(2017,52),rep(2018,52),rep(2019,52),
                           rep(2020,52))) %>%
  mutate(stateyear = paste0(State,Year) )
  

r <- r %>% 
  # Make a column to match
  mutate(stateyear = paste0(State,Year) ) %>%
  # Add in the missing states in each year
  bind_rows(stateyear %>% filter(!stateyear %in% r$stateyear)) %>%
  # Remove the stateyear
  select(-c(stateyear))


# Long format for mapping
r1 <- r %>% 
  pivot_longer(-c(State,YearID,Year,Abbrev,Code)) %>%
  left_join(states) %>%
  # Order the states
  mutate(State = factor(State,
                        levels=c("National Average",sort(unique(states$State)))))


# Read in the latest file on the missing data

missing <- read_csv("../website_ready_data/missing_data_table_combined_v3.csv")
## THIS WILL NEED DESCRIPTIVES TOO!!

```



<!-- We want tabs so we first need to have text as a heading, add the tabset feature here. Then anything else added as a heading under it will be under the tab! Arguments are .tabset-pills and .tabset-fade --> 

<br>

# EHDI DATA HUB {.tabset}
<br>

## Thoughts to move us forward

- Need to pull in descriptive terms and to order the variables! 

- We can store the annotations, terms, etc in a single Word file to read in.

- In our website plan we split raw and calculated variables. Do we want to make separate figures for them here? That sounds easier than adding another filter. Or maybe it's okay now that we know how to do it. 

## State trends over time 

- Select the measurement at the top to see how states compare over time, 2014-2020.

- Hover over a line to highlight that state. 

- Double-click a state on the right to focus on that state. Click another to compare. Double-click to reset.



```{r}

#https://plotly-r.com/client-side-linking.html

# initializing a crosstalk shared data object  
plotdat <- highlight_key(r1)  # long data here

# Filter dropdown
question_filter <- crosstalk::filter_select(
   "filter", "Select a variable to examine",
   plotdat, ~name, multiple = F
)

# Plotting:
plot <-  plot_ly(plotdat, 
    x = ~Year, y = ~value, split = ~State, text = ~name,
    # type="scatter", 
    mode = "line",
    textposition = "top"
  ) %>%
  group_by(State) %>%
  # Add in some highlighting features to see if it works and increases interactivity.
    highlight(on = "plotly_hover", off = 'plotly_doubleclick',
              selectize = FALSE, # dropdown widget
              dynamic = FALSE,  # control highlighting color
              persistent = FALSE  # retain previous selections
    )

# Found this code here: https://stackoverflow.com/questions/64234585/selecting-a-default-value-in-an-r-plotly-plot-using-a-selectize-box-via-crosstal
# Will need to see if it works or not embedded in a website...
# Just putting things together for easy display:
shiny::tags$div(class = 'flexbox',
                question_filter,
                shiny::tags$br(),
                plot)


```


## Annual map comparison

With this code we can map to a state the results of a variable in a year. Users can select a single variable and a single year.

```{r}

# In case the figures end up trying to speak with one another, we'll just make
# new objects here. 

# Establish a SharedData object to use. 
# The key here appears to be not having a key specified at all.
r1SD <- highlight_key(r1)

# Set up the widgets to be used for the mapping
widgets <- 
  # We can split the space occupied by each widget for easier viewing - total is 12
  bscols(widths = c(8, 4),
         # Create a select filter for the variable
         filter_select("variableselect", "Please select measurement here", r1SD, ~name, multiple = F),
         # Create a select filter for the year - seems the only filter to select single year
         filter_select("yearselect", "Please select year here", r1SD, ~Year, multiple = F)
  )

# Now display the connected items - embed all in the bs rather than more objects
bscols(
  widths=(12),  # max is 12 so both elements should fit the width 
  widgets,
  # Plot a map of the SharedData object
  r1SD %>%
    plot_geo(locationmode = 'USA-states') %>% # Set the location to states
    # group_by(name,Year) %>%  # I don't think we need this here
    add_trace(locations=~Code, # must be state abbreviation
              z=~value,color=~value, # we want to see the map colored by the value
              text=~c(Year)) %>%  # nice way to verify when looking at a map that the year is as intended
    layout(
      # Change the view to be specific to the US
      geo = list(scope = 'usa',
                 projection = list(type = 'albers usa'),
                 lakecolor = toRGB('white')
      )) 
)


```



I think the next step we can take here, once we have more information and also the know-how, is to include automatically populated annotations at the bottom of the figure that describes the missing data.

<br>

## Missing data

We can pull in that missing data table and format it here likely.

The table below notes the years of missing data for each state and each variable. For example, data was unavailable for Alabama in 2014 on the number of babies who failed inpatient screening and were referred straight to diagnostics (ReInStDia). 

```{r}

kable(missing %>%
        rename(Measurement=Variable),
      # Align all text to the left
      align="l")  %>%
  # Let's identify the variables a little better
  column_spec(1, bold = T) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive","bordered"),
                full_width = F)

```

## Data sources

We can document this here or in a webpage tab.

## How to use the data/site

We can document this here or in a webpage tab. 